%--------------------------------------------------------------------------------
\documentclass[]{AbstractYIC2015}

% --------------------------------------------------------------------------------
% Please do not use any extra latex packages
%--------------------------------------------------------------------------------
\usepackage{graphicx}
\usepackage{tabularx}


% --------------------------------------------------------------------------------
% Abstract's title, with capital letter only at the beginning
%--------------------------------------------------------------------------------
\title{Modeling the I/O behavior of the NEST simulator using a proxy}

% -------------------------------------------------------------------------------
% List of authors
% --------------------------------------------------------------------------------
% Put the initials and surname of the first author ('et al.' if applicable) in
% square brackets before the command \author
%
% Identify the corresponding author with the command \corref and each author
% with the command \authref{a,b,...} according to the affiliation
%
\author[T. Schumann et al.]{%
  T. Schumann\authref{a}\corref,
  W. Frings\authref{b},
  A. Peyser\authref{c},
  W. Schenck\authref{c},
  K. Thust\authref{b},
  J.M. Eppler\authref{c}
}

% --------------------------------------------------------------------------------
% Affiliations
% --------------------------------------------------------------------------------
\address{\authaddr{a}{Institute of Neuroscience and Medicine (INM-6), Computational and Systems Neuroscience \\ %
    Institute for Advanced Simulation (IAS-6) \\ %
    J\"ulich Aachen Research Alliance \\ %
    Forschungszentrum J\"ulich GmbH \\ %
    52425 J\"ulich, Germany}
  \authaddr{b}{J\"ulich Supercomputing Centre \\ %
    Institute for Advanced Simulation \\ %
    Forschungszentrum J\"ulich GmbH\\ %
    52425 J\"ulich, Germany}
  \authaddr{c}{Simulation Lab Neuroscience - Bernstein Facility for Simulation and Database Technology \\ %
    Institute for Advanced Simulation \\ %
    J\"ulich Aachen Research Alliance \\ %
    Forschungszentrum J\"ulich GmbH \\ %
    52425 J\"ulich, Germany}
}
%
% --------------------------------------------------------------------------------
% Email address of the corresponding author
% --------------------------------------------------------------------------------
\corauth{till.schumann@rwth-aachen.de}

% --------------------------------------------------------------------------------
% Keywords
% --------------------------------------------------------------------------------
\keywords{parallel I/O; simulation; neuronal networks; supercomputer;
  threading; MPI}

\abstract{
  Over the past 20 years, the NEST Initiative
  (\texttt{http://www.nest-initiative.org/}) has developed the NEST
  \cite{NEST} simulator for spiking neural network models. It is used
  in computational neuroscience to simulate the dynamics of the
  interaction between nerve cells. The systems explored with NEST
  range from small networks simulated on local machines up to large
  brain-scale circuits using the full capabilites of the world's
  leading supercomputers. To have this flexibility, NEST is
  parallelized in a hybrid fashion using threads on a compute node and
  MPI to communicate between the compute nodes \cite{Plesser07}.
  Storing simulation data from such a massively parallel application
  efficiently during runtime is essential for neuroscientic studies,
  but not a trivial task on a supercomputer with centralized storage.

  To obtain measurements from a simulation in NEST, virtual devices
  can be connected to simulated nerve cells. These devices store
  values associated with neurons by writing time to variables mappings
  such as spiking activity or membrane voltage of nerve cells to
  disk. Currently, each measurement device opens its own file using
  plain C++ streams.  With an increasing number of devices and compute
  nodes, writing data to disk in this way becomes the bottleneck for
  the whole simulation. Due to the constant growth in simulation sizes
  and therefore compute nodes, a new writing strategy has become
  necessary.

  One solution to the problem is the use of efficient parallel I/O
  libraries. It is however not clear which of the available libraries
  is best. In order to assess the characteristics of different
  libraries, we have implemented a \emph{proxy} which imitates the
  writing behavior of NEST based on a small set of parameters such as
  the number of nerve cells to record from, the number of threads and
  processes, or the rate at which data is recorded.

  This technique has allowed us to run benchmarks and carry out
  statistical analysis of the disk writing timings and optimizations
  of the interfaces without the computational costs and complexity of
  running a full NEST simulation, allowing a larger number of
  measurements for a larger number of parameter variations in less
  time. Currently, the proxy includes experimental interfaces for
  plain text, HDF5 \cite{hdf2010hierarchical} and SIONlib
  \cite{frings2009scalable} which are parametrized with data from
  profiling runs of real NEST simulations. The results of the
  benchmarks will guide our choice towards a new I/O sub-system for
  NEST.
}

\begin{document}

\maketitle



% --------------------------------------------------------------------------------
% Bibliography
% --------------------------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{frings2009scalable}
Wolfgang Frings, Felix Wolf, and Ventsislav Petkov.
\newblock Scalable massively parallel i/o to task-local files.
\newblock In {\em High Performance Computing Networking, Storage and Analysis,
  Proceedings of the Conference on}, pages 1--11. IEEE, 2009.

\bibitem{NEST}
Marc-Oliver Gewaltig and Markus Diesmann.
\newblock {NEST} ({NE}ural {S}imulation {T}ool).
\newblock {\em Scholarpedia}, 2(4):1430, 2007.

\bibitem{hdf2010hierarchical}
HDF Group 
\newblock Hierarchical data format version 5, 2000-2010.
\newblock {\em URL: http://www. hdfgroup. org/HDF5}, 2010.

\bibitem{Plesser07}
Hans~Ekkehard Plesser, Jochen~Martin Eppler, Abigail Morrison, Markus Diesmann,
  and Marc-Oliver Gewaltig.
\newblock Efficient parallel simulation of large-scale neuronal networks on
  clusters of multiprocessor computers.
\newblock In A.-M. Kermarrec, L.~Boug{\'e}, and T.~Priol, editors, {\em
  Euro-Par 2007: Parallel Processing}, volume 4641 of {\em Lecture Notes in
  Computer Science}, pages 672--681, Berlin, 2007. Springer-Verlag.

% --------------------------------------------------------------------------------
% Bibliography - the end.
% --------------------------------------------------------------------------------
\end{thebibliography}

\end{document}
